Polymethylation scores paper

### 2021-10-01 Papers/Polymethylation_scores/04_background.Rmd
started discussion outline

P2: DNA methylation could provide a reliable and specific biomarker of prenatal maternal smoke exposure. Cordblood and peripheral blood DNA methylation has been significantly associated with prenatal maternal smoking. **Gap**: Most previous studies have used blood samples, but saliva is an easier to collect tissue. To our knowledge, no study has measured associations between prenatal maternal smoking and DNA methylation in saliva samples.

P3: DNA methylation has been associated with prenatal maternal smoking at various cross-sectional time points across the life course. Ideally, a biomarker of a prenatal exposure would persist across the life course. **Gap**: Few studies have compared the effectiveness of DNA methylation as biomarker of prenatal maternal smoking in the same individuals at different times.

P4: DNA methylation has been well replicated in primarily European ancestry cohorts. The performance of other biomarkers and risk scores varies across ancestry groups. To be applied clinically without exacerbating health disparities, it is important to understand how the perfomance of biomarkers vary across ancestry groups. **Gap**: Yet few (no?) studies have compared associations of prenatal maternal smoking across diffferent ancestry groups.

P5: Choice of how to construct a DNA methylation based biomarker.

>>>>

P2: DNA methylation could provide a reliable and specific biomarker of prenatal maternal smoke exposure. Explanation of different approaches: polymethylation scores, individual CpG sites, summary methylation measures like clocks, etc.

P3: Cordblood and peripheral blood DNA methylation has been significantly associated with prenatal maternal smoking.

**Gap**: Most previous studies have used blood samples, but saliva is an easier to collect tissue. To our knowledge, no study has measured associations between prenatal maternal smoking and DNA methylation in saliva samples.

P4: DNA methylation has been associated with prenatal maternal smoking at various cross-sectional time points across the life course. Ideally, a biomarker of a prenatal exposure would persist across the life course.

**Gap**: Few studies have compared the effectiveness of DNA methylation as biomarker of prenatal maternal smoking in the same individuals at different times.

P5: DNA methylation has been well replicated in primarily European ancestry cohorts. The performance of other biomarkers and risk scores varies across ancestry groups. To be applied clinically without exacerbating health disparities, it is important to understand how the perfomance of biomarkers vary across ancestry groups.

**Gap**: Yet few (no?) studies have compared associations of prenatal maternal smoking across diffferent ancestry groups.
